---
title: "visualization_1"
author: "Jake Vettoretti"
date: "2025-09-25"
output: html_document
---

## Setup

```{r}
library(tidyverse)
library(ggridges)

library(p8105.datasets)
data("weather_df")

weather_df
```

We’ll start with a basic scatterplot to develop our understanding of ggplot’s data + mappings + geoms approach, and build quickly from there.

## Basic Scatterplot

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax))
```

We’ll move straight into defining aesthetic mappings. To create a basic scatterplot, we need to map variables to the X and Y coordinate aesthetics:

Well, my “scatterplot” is blank. That’s because I’ve defined the data and the aesthetic mappings, but haven’t added any geoms: ggplot knows what data I want to plot and how I want to map variables, but not what I want to show. Below I add a geom to define my first scatterplot.

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```

If want to pre-process with piping:

The code below could be used instead to produce the same figure. Using this style can be helpful if you want to do some pre-processing before making your plot but don’t want to save the intermediate data. It’s also consistent with many other pipelines: you start with a data frame, and then do stuff by piping the dataframe into the next function. Most of my plotting code is written like this.

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```





